# Styling Plots {#sec-plotstyle}

```{r, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)
```

```{r}
library(tidyverse)
library(patchwork)
library(ggthemes)
```


## Aesthetics

### Colour/Fill

The `colour` argument changes the point and line colour, while the `fill` argument changes the interior colour of shapes. Type `colours()` into the console to see a list of all the named colours in R. Alternatively, you can use hexadecimal colours like `"#FF8000"` or the `rgb()` function to set red, green, and blue values on a scale from 0 to 1.

Hover over a colour to see its R name.

<style>
  .colordemo { list-item-style: none; padding: 0; margin: 0;}
  
  .colordemo li, main ul.colordemo > li:first-child { 
    display: inline-block; 
    border: 0.25px solid black; 
    margin: 0px; 
    padding: 1.3em 0.2em 0.2em 1.2em; 
  }
  
  .colordemo li span { 
    display: none; 
    background-color: white; 
    border: 0.25px solid black; 
    padding: 0.25em 0.5em; 
    border-radius: 0 1em 1em 1em;
  }
  .colordemo li:hover span { 
    display: inline; 
    position: absolute;
  }
</style>

```{r colour-demo, echo = FALSE, results='asis'}

name <- colours(TRUE)
rgb <- col2rgb(name) 
hsv <- rgb2hsv(rgb)
hex <- as.hexmode(rgb) %>% 
  as.character(keepStr = TRUE)
hex[nchar(hex) == 1] <- paste0("0", hex[nchar(hex) == 1])
hex <- apply(hex, 2, paste, collapse = "")

colors <- tibble(name = name,
       hue = hsv[1, ] %>% round(2),
       sat = hsv[2, ] %>% round(2),
       val = hsv[3, ] %>% round(2),
       hex = hex) %>%
  arrange(hue, sat, val)

glue::glue("    <li style='background-color: #{colors$hex};'><span>{colors$name}</span></li>") %>%
  paste(collapse = "\n") %>%
  paste0("\n<ul class='colordemo'>\n", ., "\n</ul>\n") %>%
  cat()
```

### Alpha

The `alpha` argument changes transparency (0 = totally transparent, 1 = totally opaque). 

```{r}
#| label: fig-alpha-demo
#| echo: false
#| fig-height: 3
#| fig-width: 10
#| fig-cap: Varying alpha values
#| fig-alt: A demo with 'alpha' on the x-axis and ticks from 0.0 to 1.0 in 0.1 steps. The top line shows blue dota with increasing alpha (i.e., decreacing transparency) over a black line. The bottom line shows black dots with increasing alpha over a blue line.
alphas <- crossing(alpha = seq(0, 1, .1), y = 0:1)
ggplot(alphas, aes(x = alpha, y = y, alpha = I(alpha), 
                   color = as.factor(y))) +
  geom_hline(yintercept = 0, size = 5, color = "dodgerblue") +
  geom_hline(yintercept = 1, size = 5, color = "black") +
  geom_point(size = 20, show.legend = FALSE) +
  scale_x_continuous("alpha", breaks = alphas$alpha) +
  scale_y_continuous("", breaks = -1:2, limits = c(-1, 2)) +
  scale_color_manual(values = c("black", "dodgerblue")) +
  theme_minimal(base_size = 20) +
  theme(axis.text.y = element_blank(),
        panel.grid.minor = element_blank())
```

### Shape

The `shape` argument changes the shape of points. 

```{r}
#| label: fig-shape-demo
#| echo: false
#| fig-height: 2
#| fig-width: 10
#| fig-cap: The 25 shape values
#| fig-alt: "The numbers 1-25 on the x-axis, with the 25 different point shapes above them: open cirlce, open up triangle, +, x, open diamond, open down triangle, box with x, *, diamond with +, circle with +, star of david, box with +, circle with x, box with up triangle, solid square, solid circle, solid up triangle, solid diamond, big solid circle, small solid circle, open circle, open square, open diamond, open up triangle, open down triangle"
shapes <- tibble(shape = 1:25, y = 0)
ggplot(shapes, aes(x = shape, y = y, shape = I(shape))) +
  geom_point(size = 8) +
  scale_x_continuous("", breaks = shapes$shape) +
  scale_y_continuous("", breaks = 0) +
  theme_minimal(base_size = 20) +
  theme(axis.text.y = element_blank(),
        panel.grid.minor = element_blank())
```

### Linetype

You can probably guess what the `linetype` argument does.


```{r}
#| label: fig-linetype-demo
#| echo: false
#| fig-height: 4
#| fig-width: 10
#| fig-cap: The 6 linetype values at different sizes
#| fig-alt: The x-axis shows linetype, with 6 values (1-6). There are 4 facets of different linewidths (0.25, 0.5, 1, 2). Linetype 1 is solid, 2 is dashes, 3 is dots, 4 is alternating dot dash, 5 is long dashes, and 6 is longer dot dash.
lines <- crossing(linetype = 1:6, linewidth = c(0.25, 0.5, 1, 2)) %>% 
  mutate(x = linetype)
ggplot(lines) +
  geom_vline(aes(xintercept = linetype, 
                 linetype = I(linetype),
                 linewidth = I(linewidth))) +
  scale_x_continuous("linetype", breaks = 1:6, 
                     expand = expansion(add = 1)) +
  facet_wrap(~linewidth, nrow = 1, labeller = label_both) +
  theme_minimal(base_size = 20) +
  theme(axis.text.y = element_blank(),
        panel.grid.minor = element_blank())
```

## Palettes

```{r}
#} include: false
palette_demo <- function(levels, seed = 8675309) {
  set.seed(seed)
  
  dat <- faux::sim_design(
    within = list(level = as.character(1:levels)),
    between = list(condition= c("control", "experimental")),
    n = 50, dv = "score",
    plot = FALSE, long = TRUE)
  
  ggplot(dat, aes(x = level, 
                  y = score, 
                  color = level, 
                  fill = level)) + 
    geom_violin(alpha = 0.8, color = "black", na.rm = TRUE) +
    stat_summary(fun = mean, na.rm = TRUE) +
    scale_x_discrete(name = "") +
    guides(color = "none", fill = "none") +
     theme(axis.text.y = element_blank(),
           axis.title.y = element_blank(),
           axis.ticks.y = element_blank())
}

palette_plot <- function(g, p) { 
  g + scale_color_brewer(palette = p) +
    scale_fill_brewer(palette = p) +
    ggtitle(glue::glue("palette = \"{p}\""))
}

viridis_d_plot <- function(g, option = "viridis") { 
  g + scale_color_viridis_d(option = option) +
    scale_fill_viridis_d(option = option) +
    ggtitle(glue::glue("option = \"{option}\""))
}


```

Discrete palettes change depending on the number of categories. 

```{r}
#| label: fig-default-discrete-palette
#| fig-width: 8
#| fig-height: 10
#| fig-cap: "Default discrete palette with different numbers of levels."
#| fig-alt: A series of violin plots showing 1 to 8 violins and the colours that are used for each combination. 
plots <- lapply(1:8, palette_demo)
wrap_plots(plots, ncol = 2)
```


### Viridis Palettes

Viridis palettes are very good for colourblind-safe and greyscale-safe plots. The work with any number of categories, but are best for larger numbers of categories or continuous colours.

#### Discrete Viridis Palettes

Set `r glossary("discrete")` viridis colours with `scale_colour_viridis_d()` or `scale_fill_viridis_d()` and set the `option` argument to one of the options below. Set `direction = -1` to reverse the order of colours.


```{r}
#| label: fig-viridis-palette
#| fig-width: 8
#| fig-height: 10
#| fig-cap: Discrete viridis palettes
#| fig-alt: "5 rows and 2 columns of violin plots. Each row shows a different colour palette from the viridis options: magma, inferno, plasma, viridis, and cividis. The left plot in each row shows the colours for 5 levels, the right plot shows the colours for 15 levles."
options <- list("magma", "inferno", "plasma", "viridis", "cividis")

g <- palette_demo(5)
plots5 <- lapply(options, viridis_d_plot, g = g)

g <- palette_demo(15)
plots15 <- lapply(options, viridis_d_plot, g = g)

( wrap_plots(plots5, ncol = 1) | 
  wrap_plots(plots15, ncol = 1)) +
  plot_layout(widths = c(1, 3))
```


::: {.callout-note}
If the end colour is too light for your plot or the start colour too dark, you can set the `begin` and `end` arguments to values between 0 and 1, such as `scale_colour_viridis_c(begin = .1, end = .9)`.
:::

#### Continuous Viridis Palettes

Set `r glossary("continuous")` viridis colours with `scale_colour_viridis_c()` or `scale_fill_viridis_c()` and set the `option` argument to one of the options below. Set `direction = -1` to reverse the order of colours.

```{r}
#| label: fig-viridis-continuous
#| echo: false
#| fig-width: 8
#| fig-height: 8
#| fig-cap: Continuous viridis palettes.
#| fig-alt: "Five hex plots showing density across the x and y axes, where lighter colours correspond to higher count.  Each panel shows a different colour palette from the viridis options: magma, inferno, plasma, viridis, and cividis."
options <- list("magma", "inferno", "plasma", "viridis", "cividis")

set.seed(8675309)
g <- faux::rnorm_multi(
    n = 10000,
    vars = 2, 
    r = 0.5) %>%
  ggplot(aes(X1, X2)) + 
  theme(axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks.y = element_blank(), 
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank())

# install hexbin if this fails
viridis_c_plot <- function(g, option, geom = geom_hex) {
  g + scale_color_viridis_c(option = option) +
    geom_hex() +
    scale_fill_viridis_c(option = option) +
    ggtitle(glue::glue("option = \"{option}\""))
}

hex_plots <- lapply(options, viridis_c_plot, g = g)
wrap_plots(hex_plots, nrow = 3)
    
```

### Brewer Palettes

Brewer palettes give you a lot of control over plot colour and fill. You set them with `scale_color_brewer()` or `scale_fill_brewer()` and set the `palette` argument to one of the palettes below. Set `direction = -1` to reverse the order of colours.

#### Qualitative Brewer Palettes

These palettes are good for `r glossary("categorical")` data with up to 8 categories (some palettes can handle up to 12). The "Paired" palette is useful if your categories are arranged in pairs.

```{r}
#| label: fig-quali-brewer
#| echo: false
#| warning: false
#| fig-width: 8
#| fig-height: 8
#| fig-cap: "Qualitative brewer palettes."
#| fig-alt: "8 violin plots showing the 8-12 colours in each palette (numebr of colours in parentheses): Accent (8), Dark2 (8), Paired (12), Pastel1 (8), Pastel2 (8), Set1 (9), Set2 (8), Set3 (12)"
palettes <- list("Accent", "Dark2", "Paired", "Pastel1", "Pastel2", "Set1", "Set2", "Set3")

g <- palette_demo(12)
plots <- lapply(palettes, palette_plot, g = g)
wrap_plots(plots, ncol = 2)
```

#### Sequential Brewer Palettes

These palettes are good for up to 9 `r glossary("ordinal")` categories with a lot of categories.

```{r}
#| label: fig-sequential-brewer
#| echo: false
#| fig-width: 8
#| fig-height: 18
#| fig-cap: "Sequential brewer palettes."
#| fig-alt: "18 violin plots showing the 9 colours in each palette: Reds, RdPu, Oranges, OrRd, YlOrRd, YlOrBr, YlGn, YlGnBu, Greens, GnBu, Blues, BuGn, BuPu, Purples, PuRd, PuBu, PuBuGn, Greys"
palettes <- c("Reds", "RdPu",
              "Oranges", "OrRd", 
              "YlOrRd", "YlOrBr", "YlGn", "YlGnBu",
              "Greens", "GnBu",
              "Blues", "BuGn", "BuPu", 
              "Purples", "PuRd", "PuBu", "PuBuGn",
              "Greys")

g <- palette_demo(9)
plots <- lapply(palettes, palette_plot, g = g)
wrap_plots(plots, ncol = 2)
```

#### Diverging Brewer Palettes

These palettes are good for `r glossary("ordinal")` categories with up to 11 levels where the centre level is a neutral or baseline category and the levels above and below it differ in an important way, such as agree versus disagree options.

```{r}
#| label: fig-diverging-brewer
#| echo: false
#| fig-width: 8
#| fig-height: 10
#| fig-cap: "Diverging brewer palettes."
#| fig-alt: "Nine violin plots showing the 11 colours in each palette: BrBG, PiYG, PRGn, PuOr, RdBu, RdGy, RdYlBu, RdYlGn, and Spectral"
palettes <- c("BrBG", "PiYG", "PRGn", "PuOr", "RdBu", "RdGy", "RdYlBu", "RdYlGn", "Spectral")

g <- palette_demo(11)
plots <- lapply(palettes, palette_plot, g = g)
wrap_plots(plots, ncol = 2)
```

## Themes {#sec-themes-appendix}

<pkg>ggplot2</pkg> has 8 built-in themes that you can add to a plot like `plot + theme_bw()` or set as the default theme at the top of your script like `theme_set(theme_bw())`.

```{r}
#| label: fig-ggplot2-themes
#| echo: false
#| fig-width: 8
#| fig-height: 8
#| fig-cap: "{ggplot2} themes."
#| fig-alt: "The same plot in 8 different themes: bw, classic, dark, gray, light, linedraw, minimal, void"
g <- palette_demo(2) + facet_wrap(~condition)

themes <- c("theme_bw", "theme_classic", "theme_dark", "theme_gray", "theme_light", "theme_linedraw", "theme_minimal", "theme_void")

plots <- lapply(themes, function(f) {
  g + get(f)() + ggtitle(glue::glue("{f}()"))
})
wrap_plots(plots, ncol = 2)
```

### ggthemes

You can get more themes from add-on packages, like <pkg>ggthemes", "https://yutannihilation.github.io/allYourFigureAreBelongToUs/ggthemes/")`. Most of the themes also have custom `scale_` functions like `scale_colour_economist()`. Their website has extensive examples and instructions for alternate or dark versions of these themes.

```{r}
#| label: fig-ggthemes
#| echo: false
#| fig-width: 8
#| fig-height: 18
#| fig-cap: "{ggthemes} themes."
#| fig-alt: "The same plot in 17 different themes: calc, economist, economist_white, excel, excel_new, few, fivethirtyeight, gdocs, hc, igray, map, pander, solarized, solarized_2, stata, tufte, and wsj."
g <- palette_demo(2) + facet_wrap(~condition)

themes <- c("theme_wsj", "theme_tufte", "theme_stata", "theme_gdocs", "theme_fivethirtyeight", "theme_economist", "theme_igray", "theme_hc", "theme_few", "theme_map", "theme_excel", "theme_calc", "theme_economist_white", "theme_excel_new", "theme_pander", "theme_solarized", "theme_solarized_2") %>% sort() %>% setNames(., .)

plots <- lapply(themes, function(f) {
  scale_colour <- tryCatch({
    get(gsub("theme_","scale_colour_", gsub("_(2|white)", "", f)))()
  }, error = function(e) {
    return(NULL)
  })
  
  scale_fill <- tryCatch({
    get(gsub("theme_","scale_fill_", gsub("_(2|white)", "", f)))()
  }, error = function(e) {
    return(NULL)
  })
  
  g + get(f)() + 
    ggtitle(glue::glue("{f}()")) + 
    scale_colour + 
    scale_fill
})

wrap_plots(plots, ncol = 2)
```


### Fonts

You can customise the fonts used in themes. All computers should be able to recognise the families "sans", "serif", and "mono", and some computers will be able to access other installed fonts by name.

```{r}
#| include: false
if (.Platform$OS.type == "windows") {
  windowsFonts("Comic Sans MS" = windowsFont("Comic Sans MS"))
}
```

```{r}
#| label: fig-theme-font-demo
#| fig-width: 10
#| fig-height: 2.5
#| fig-cap: Different fonts.
#| fig-alt: The same plot with the stystem standard Sans, Serif, Mono, fonts, plus Comic Sans MS.
sans <- g + theme_bw(base_family = "sans") + 
  ggtitle("Sans")
serif <- g + theme_bw(base_family = "serif") + 
  ggtitle("Serif")
mono <- g + theme_bw(base_family = "mono") + 
  ggtitle("Mono")
font <- g + theme_bw(base_family = "Comic Sans MS") + 
  ggtitle("Comic Sans MS")

sans + serif + mono + font + plot_layout(nrow = 1)
```


::: {.callout-warning}
If you are working on a Windows machine and get the error "font family not found in Windows font database", you may need to explicitly map the fonts. In your setup code chunk, add the following code, which should fix the error. You may need to do this for any fonts that you specify.

```{r}
#| eval: false
windowsFonts("Comic Sans MS" = windowsFont("Comic Sans MS"))
```
:::


The <pkg>showtext</pkg> package is a flexible way to add fonts.

If you have a .ttf file from a font site, like [Font Squirrel](https://www.fontsquirrel.com), you can load the file directly using `font_add()`. Set `regular` as the path to the file for the regular version of the font, and optionally add other versions. Set the `family` to the name you want to use for the font. You will need to include any local font files if you are sharing your script with others.

```{r}
library(showtext)

# font from https://www.fontsquirrel.com/fonts/SF-Cartoonist-Hand

font_add(
  regular = "fonts/cartoonist/SF_Cartoonist_Hand.ttf",
  bold = "fonts/cartoonist/SF_Cartoonist_Hand_Bold.ttf",
  italic = "fonts/cartoonist/SF_Cartoonist_Hand_Italic.ttf",
  bolditalic = "fonts/cartoonist/SF_Cartoonist_Hand_Bold_Italic.ttf",
  family = "cartoonist" 
)
```

To download fonts directly from [Google fonts](https://fonts.google.com/), use the function `font_add_google()`, set the `name` to the exact name from the site, and the `family` to the name you want to use for the font.

```{r}
# download fonts from Google
font_add_google(name = "Courgette", family = "courgette")
font_add_google(name = "Poiret One", family = "poiret")
```

After you've added fonts from local files or Google, you need to make them available to R using `showtext_auto()`. You will have to do these steps in each script where you want to use the custom fonts.

```{r}
showtext_auto() # load the fonts
```

To change the fonts used overall in a plot, use the `theme()` function and set `text` to `element_text(family = "new_font_family")`.

```{r}
#| label: fig-font-demo
#| fig-width: 8
#| fig-height: 2.5
#| fig-cap: Custom Fonts.
#| fig-alt: "The same plot with three fonts: Courgette (italic handwriting), Cartoonist Hand (cartoon style), and Poiret One (thin art deco)."
a <- g + theme(text = element_text(family = "courgette")) +
  ggtitle("Courgette")
b <- g + theme(text = element_text(family = "cartoonist")) +
  ggtitle("Cartoonist Hand")
c <- g + theme(text = element_text(family = "poiret")) +
  ggtitle("Poiret One")

a + b + c
```

To set the fonts for individual elements in the plot, you need to find the specific argument for that element. You can use the argument `face` to choose "bold", "italic",  or "bolditalic" versions, if they are available.

```{r}
#| label: fig-demo-multi-text
#| fig-width: 6
#| fig-height: 2.5
#| fig-cap: Multiple custom fonts on the same plot.
#| fig-alt: The same plot with the title in Cartoonist Hand bold, the facet strip titles in cartoonist hand italic, and the x-axis labels in sans serif font.
g + ggtitle("Cartoonist Hand") +
  theme(
    title = element_text(family = "cartoonist", face = "bold"),
    strip.text = element_text(family = "cartoonist", face = "italic"),
    axis.text = element_text(family = "sans")
  )
```

