# Accessibility {#sec-accessibility}

```{r}
#| label: setup-accessibility
#| message: false

library(tidyverse)
library(sonify)
```

## Visual Impairment

Here is some code to simulate different relationships.

```{r}
dat <- tibble(
  x = seq(-3, 3, .01),
) |>
  mutate(
    none = rnorm(n()),
    linear_pos = x + rnorm(n()),
    linear_neg = -x + rnorm(n()),
    quadratic = x^2 + rnorm(n()),
    category = sample(c("control", "experimental"), n(), TRUE)
  )
```


### Sonify

@siegert2017sonify

No relationship sounds like a steady or randomly wavering tone.

```{r}
#| label: fig-none
ggplot(dat, aes(x = x, y = none)) +
  geom_point() +
  geom_smooth()
```

```{r}
sonify(dat$x, dat$none)
```


A linear positive relationship sounds like an increasing tone.

```{r}
#| label: fig-linear-pos
ggplot(dat, aes(x = x, y = linear_pos)) +
  geom_point() +
  geom_smooth()
```

```{r}
sonify(dat$x, dat$linear_pos)
```

A linear negative relationship sounds like a decreasing tone.

```{r}
#| label: fig-linear-neg
ggplot(dat, aes(x = x, y = linear_neg)) +
  geom_point() +
  geom_smooth()
```

```{r}
sonify(dat$x, dat$linear_neg)
```

This quadratic relationship decreases in pitch and then increases.

```{r}
#| label: fig-quadratic
ggplot(dat, aes(x = x, y = quadratic)) +
  geom_point() +
  geom_smooth()
```

```{r}
sonify(dat$x, dat$quadratic)
```

### BrailleR

[BrailleR](https://ajrgodfrey.github.io/BrailleR/) @godfrey2020brailler

```{r}
#| message: false

# BrailleR was removed from CRAN recently, if it's not back, try
# remotes::install_github("ajrgodfrey/BrailleR")
library(BrailleR)
```

```{r}
ggplot(dat, aes(x = x, y = linear_neg)) +
  geom_point() +
  geom_smooth()
```


```{r}
#| label: fig-violin
ggplot(dat, aes(x = category, y = quadratic)) +
  geom_violin()
```

## Resources

* [arXiv Accessibility Forum 2023](https://accessibility2023.arxiv.org/index.html)
* [Joo Young Accessible Data Science for the Blind Using R](https://jooyoungseo.me/post/ds4blind/)
* [Resources For Using R With Screen Readers](https://ropensci.org/blog/2024/09/05/screen-readers-tools/)
