# Conventions {#sec-conventions}

This book will use the following conventions:

* Code: `list(number = 1, letter = "A")`
* File paths: `r path("data/sales.csv")`
* Packages: <pkg>tidyverse</pkg>
* Menu/interface options: <if>Tools > Global Options... > Pane Layout</if>
* Glossary items: `r glossary("ordinal")`
* Citations: @R-tidyverse
* Internal links: @sec-intro
* External links: [Mastering Shiny](https://mastering-shiny.org/)
* Mac-specific: <mac>Cmd-Shift-F10</mac> 
* Windows-specific: <pc>Ctl-Shift-F10</pc>

A list of mac and windows [keyboard shortcuts](https://support.posit.co/hc/en-us/articles/200711853-Keyboard-Shortcuts-in-the-RStudio-IDE).


## Test Yourself

I am going to learn a lot: `r torf(TRUE)`

```{r, echo=FALSE, results='asis'}
opts <- c(
   "the probability that the null hypothesis is true",
   answer = "the probability of the observed, or more extreme, data, under the assumption that the null-hypothesis is true",
   "the probability of making an error in your conclusion"
)

cat("* What is a p-value?", longmcq(opts))
```


```{block, webex.hide="Hidden Solutions"}
You found it!
```

## Callout boxes

[See the [quarto reference](https://quarto.org/docs/authoring/callouts.html) for more options.]{.aside}

::: {.callout-note}
Informational asides.
:::

::: {.callout-tip}
Tips
:::

::: {.callout-warning}
Notes to warn you about something.
:::

::: {.callout-caution}
Notes about things that could cause serious errors.
:::

::: {.callout-important}
Notes about things that are important.
:::

::: {.callout-note .try collapse="true"}
## Try (click to expand)
Try it yourself.
:::

## Code and Output

```{verbatim, lang = "markdown"}
## Markdown Example

* Inline code: `r nrow(iris)`
* *Italics*
* **Bold**
```

```{r}
# code chunks
paste("Code", "Output", 1, sep = " ")
```


```{r}
#| echo: fenced
#| fig-cap: "Set `#| echo: fenced` for code chunks with headers"
#| label: fig-fenced-example
#‎| fig-cap: "#| Set `echo: fenced` for code chunks with headers"
#‎| echo: false

hist(rnorm(100000))
```

:::{.callout-warning collapse="true"}
## Showing `echo` in fenced code blocks

Use a version of `#‎|` with an invisible character to show options in a fenced code block that you don't actually want to apply to this chunk, such as a second `echo` or a duplicate `label`. Copy and edit the text below into RStudio and you'll see a red highlighted dot between `#` and `|`. These lines have to go *below* any lines that you do want to apply to this code block.

```
#‎| echo: false
```
:::





