# Assessment 1

```{r, include=FALSE}
library(tidyverse)
library(gt)
```


A project with a qmd file showing basic reproducible report skills

## Instructions

Your task is to write the outline structure of your thesis report in a reproducible quarto script. The report must contain the following:

* [ ] The main section headers that you expect to use in your report (e.g., Abstract, Introduction, Methods, Results, Discussion, References) and any subheaders that might be relevant.
* [ ]  A short *paragraph* about your research topic in the Introduction section 
* [ ] One relevant *image* with a figure caption that is references in the text
* [ ] A *table* in the Methods section that gives a rough timetable for your thesis, created using markdown or data frames in R
* [ ] At least two *bibliographic references* cited in the text, created and displayed in the reference section using reproducible techniques


## Submission

* Due: Wednesday, 2024-10-09 by noon
* Covers: chapters 1 and 2
* Worth: 10%
* Do not put your name in your report; use your student ID as the author.
* Please submit a **zip file** containing: 
  1. the .rproj file
  2. your reproducible script, named `report1_studentID.qmd`
  3. any additional files necessary to reproduce your report(e.g., images or bibliography files),
  4. the rendered html report, named `report1_studentID.html`.

## Marking Rubric


```{r}
rubric <- list(
  "Research & Knowledge" = list(
    "**Topic1**: Description" = list(
      "A-Excellent" = "",
      "B-Very Good" = "",
      "C-Good" = "",
      "D-Satisfactory" = "",
      "E-Poor" = ""
    )
  ),
  "Evaluation" = list(
    "**Topic2**: Description" = list(
      "A-Excellent" = "",
      "B-Very Good" = "",
      "C-Good" = "",
      "D-Satisfactory" = "",
      "E-Poor" = ""
    )
  ),
  "Communication" = list(
    "**Topic3**: Description" = list(
      "A-Excellent" = "",
      "B-Very Good" = "",
      "C-Good" = "",
      "D-Satisfactory" = "",
      "E-Poor" = ""
    )
  )
)
```

```{r}
colnames <- c("ILO", 
              "A: Excellent", 
              "B: Very Good", 
              "C: Good", 
              "D: Satisfactory", 
              "E: Poor")
cats <- rep(names(rubric), sapply(rubric, length))

reduce(rubric, c) |>
  as_tibble() |>
  t() |>
  as_tibble(rownames = "ILO", .name_repair = "minimal") |>
  `names<-`(colnames) |>
  mutate(Category = cats) |>
  select(Category, everything()) |>
  group_by(Category) |>
  gt::gt() |>
  gt::fmt_markdown(columns = "ILO") |>
  gt::tab_style(style = list(gt::cell_text(align = "left", v_align= "top")),
                locations = cells_body()) |>
  gt::tab_style(style = list(gt::cell_fill("#4D9FCF66")),
                locations = cells_row_groups()) |>
  gt::tab_style(style = list(gt::cell_borders("bottom", "grey")),
                locations = cells_body()) |>
  gt::opt_stylize(style = 6, add_row_striping = FALSE)

```

