# Assessment 3

```{r, include=FALSE}
library(tidyverse)
library(gt)
```

A project demonstrating data wrangling skills

## Instructions

TBA

## Submission

* Covers: chapters 1-7, emphasising 5-7
* Worth: 30%
* Do not put your name in your report; use your student ID as the author.
* Please submit a **zip file** containing: 
  1. the .rproj file
  2. your reproducible script, named `report3_studentID.qmd`
  3. any additional files necessary to reproduce your report(e.g., images or bibliography files),
  4. the rendered html report, named `report3_studentID.html`.

## Marking Rubric


```{r}
#| echo: false
rubric <- list(
  "Research & Knowledge" = list(
    "**Topic1**: Description" = list(
      "A-Excellent" = "",
      "B-Very Good" = "",
      "C-Good" = "",
      "D-Satisfactory" = "",
      "E-Poor" = ""
    )
  ),
  "Evaluation" = list(
    "**Topic2**: Description" = list(
      "A-Excellent" = "",
      "B-Very Good" = "",
      "C-Good" = "",
      "D-Satisfactory" = "",
      "E-Poor" = ""
    )
  ),
  "Communication" = list(
    "**Topic3**: Description" = list(
      "A-Excellent" = "",
      "B-Very Good" = "",
      "C-Good" = "",
      "D-Satisfactory" = "",
      "E-Poor" = ""
    )
  )
)
```

```{r}
#| echo: false
colnames <- c("ILO", 
              "A: Excellent", 
              "B: Very Good", 
              "C: Good", 
              "D: Satisfactory", 
              "E: Poor")
cats <- rep(names(rubric), sapply(rubric, length))

reduce(rubric, c) |>
  as_tibble() |>
  t() |>
  as_tibble(rownames = "ILO", .name_repair = "minimal") |>
  `names<-`(colnames) |>
  mutate(Category = cats) |>
  select(Category, everything()) |>
  group_by(Category) |>
  gt::gt() |>
  gt::fmt_markdown(columns = "ILO") |>
  gt::tab_style(style = list(gt::cell_text(align = "left", v_align= "top")),
                locations = cells_body()) |>
  gt::tab_style(style = list(gt::cell_fill("#4D9FCF66")),
                locations = cells_row_groups()) |>
  gt::tab_style(style = list(gt::cell_borders("bottom", "grey")),
                locations = cells_body()) |>
  gt::opt_stylize(style = 6, add_row_striping = FALSE)

```


